---
title: "Transfer data"
author: "Nathan Matten"
date: "3/27/2022"
output:
  html_document: 
    number_sections: yes
    toc: yes
    toc_float: yes
    fig_caption: yes
    fig_width: 6
    fig_height: 4
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(shipunov)
library(caret)
set.seed(2141122)
```

We're only going to look at players that transfer into the top 6 conferences
Loading in data
```{r}
#Loading players for each conference they transferred into
acc_2017 <- read.csv(file = "ACC_2017_22.csv", header = T)
big12_2017 <- read.csv(file = "Big12_2017_22.csv", header = T)
big10_2017 <- read.csv(file = "Big10_2017_22.csv", header = T)
bigeast_2017 <- read.csv(file = "BigEast_2017_22.csv", header = T)
sec_2017 <- read.csv(file = "SEC_2017_22.csv", header = T)
pac12_2017 <- read.csv(file = "Pac12_2017_22.csv", header = T)
tran_2015 <- read.csv(file = "trank_data_15.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2015$player = paste(tran_2015$player_name, "(15)")
tran_2016 <- read.csv(file = "trank_data_16.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2016$player = paste(tran_2016$player_name, "(16)")
tran_2017 <- read.csv(file = "trank_data_17.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2017$player = paste(tran_2017$player_name, "(17)")
tran_2018 <- read.csv(file = "trank_data_18.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2018$player = paste(tran_2018$player_name, "(18)")
tran_2019 <- read.csv(file = "trank_data_19.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2019$player = paste(tran_2019$player_name, "(19)")
tran_2020 <- read.csv(file = "trank_data_20.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2020$player = paste(tran_2020$player_name, "(20)")
tran_2021 <- read.csv(file = "trank_data_21.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2021$player = paste(tran_2021$player_name, "(21)")
tran_2022 <- read.csv(file = "trank_data_22.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2022$player = paste(tran_2022$player_name, "(22)")
cbb_teams <- read.csv(file = "cbb_teams.csv", header = T)
# Creating mega reference table
tran_stats <- rbind(tran_2015,tran_2016,tran_2017,tran_2018,tran_2019,tran_2020,tran_2021,tran_2022) %>% 
  select(player,GP,Min_per,ORtg,usg,eFG,TS_per,ORB_per,DRB_per,AST_per,TO_per,FTM,FTA,FT_per,twoPM,twoPA,twoP_per,TPM,TPA,TP_per,blk_per,stl_per,ftr,yr,porpag,pid,BPM,year)
``` 

```{r}
# Making 1 giant transfer file for the big 6 conferences
tran <- rbind(big10_2017,big12_2017,bigeast_2017,acc_2017,sec_2017,pac12_2017) %>% 
  select(-RK) %>% 
  arrange(PLAYER)
# Getting conferences for each transfer (from and to)
tran_team <- 
  tibble(
    player = tran$PLAYER,
    year = tran$YEAR.1,
    old_team = tran$OLD.TEAM,
    new_team = tran$TEAM,
    grade = tran$YEAR,
   ) %>% 
  arrange(old_team, year)

# Mapping new teams
tran_team <-  merge(tran_team, cbb_teams, by.x = "old_team", by.y = "Team", all.x = TRUE) 
tran_team <-  merge(tran_team, cbb_teams, by.x = "new_team", by.y = "Team", all.x = TRUE, no.dups = TRUE) 

# Setting up and arranging new table
tran_team <- tran_team %>%
  mutate(old_conference = Conference.x) %>% 
  mutate(new_conference = Conference.y) %>% 
  select(player,old_team,old_conference,new_team,new_conference,grade,year) %>%
  # Used this to search for conferences that didn't match in the mapping file
  #filter(is.na(old_conference)) %>% 
  arrange(old_conference,old_team,year)
# adding in conference ratings
tran_team$conf_rating <- 
ifelse (tran_team$old_conference %in% c('ACC','SEC','Pac-12','Big 10','Big 12','Big East'),1,
  ifelse (tran_team$old_conference %in% c("AAC","MWC","Atlantic 10","WCC","MVC"),2,
    ifelse (tran_team$old_conference %in% c("MAC","C-USA","SoCon","CAA","Sun Belt","Ivy League"),3,
      ifelse (tran_team$old_conference %in% c("Summit League","WAC","Horizon","Big West","Big Sky","MAAC","OVC"),4,
        ifelse (tran_team$old_conference %in% c("Patriot","ASUN","Big South","America East"),5,
          ifelse (tran_team$old_conference %in% c("Southland","NEC"),6,7)
        )
      )
    )
  )
)
# Creating the large table of transfers and their stats based off the two tables (tran and tran_team above)
transfer <- merge(tran_team, tran, by.y = "PLAYER", by.x = "player", all.x = TRUE) %>% 
  arrange(old_conference,year,old_team,new_conference,new_team)
# Middle table to get P-ID to then go back and remap
t2 <- merge(tran_team, tran_stats, by.x = "player", by.y = "player") %>% 
  select(player,old_team,old_conference,new_team,new_conference,grade,year.x,pid,conf_rating) %>% 
  arrange(year.x,pid)
# Mapping based on P-ID... then creating the mega-table of desired data
t3 <- merge(t2,tran_stats, by.x = "pid", by.y = "pid") %>% 
  mutate(player = player.x) %>%
  mutate(tran_year = year.x) %>% 
  select(-player.y,-player.x,-year.x) %>% 
  arrange(old_conference,old_team,new_conference,new_team,player,year)
# Figuring out how to only pull the transfer year stats and the previously played year's stats
#tran_final <- tran_yay %>% 
#  filter((year == tran_year))# & row((year == tran_year)-1))
```
Creating final tables to use for analysis
```{r}
#Using this as testing ground to get rows needed for data
# Rows from season before transfer
rows <- which(t3$year[] == t3$tran_year[])
# Rows from season after transfer
nrows <- rows - 1
# Finding all years that contain players that transferred
frows <- c(rows,nrows)
# Creating a tibble from data before players transferred
pre_t <- t3 %>% 
  filter(row_number() %in% nrows) %>% 
  select(pid,player,year,tran_year,old_conference,old_team,new_conference,new_team,BPM,conf_rating,usg,eFG,Min_per,TS_per,TP_per,blk_per,stl_per) %>% 
  arrange(old_conference,old_team,new_conference,new_team,player,year)
# Creating a tibble from data after players transferred, arranged the same as above
post_t <- t3 %>% 
  filter(row_number() %in% rows) %>% 
  select(pid,player,year,tran_year,old_conference,old_team,new_conference,new_team,BPM,conf_rating,usg,eFG,Min_per,TS_per,TP_per,blk_per,stl_per) %>% 
  rename(BPM2 = BPM,
         conf_rating2 = conf_rating,
         usg2 = usg,
         eFG2 = eFG,
         Min_per2 = Min_per,
         TS_per2 = TS_per,
         TP_per2 = TP_per,
         blk_per2 = blk_per,
         stl_per2 = stl_per) %>% 
  arrange(old_conference,old_team,new_conference,new_team,player,year)
# Combining data based on row, so we can do further analysis (as needed)
tran_final <- cbind(pre_t, post_t)
# Creating a binary value whether a player is BPM positive or negative each year
tran_final$BPMbin <- ifelse(tran_final$BPM >= 0, 1, 0)
tran_final$BPMbin2 <- ifelse(tran_final$BPM2 >= 0, 1, 0)
tran_final$year <- as.numeric(tran_final$year)
```

```{r} # IGNORE THIS CHUNK 
# Now running KNN Bootstrap... let's see how it goes
#tran_sub <- which(tran_final$year != tran_final$tran_year)
#boots <- tran_final %>% 
#  select(year,BPMbin,conf_rating)
#kboots <- BootKNN(boots,classes = tran_final$BPMbin,sub = tran_sub, nboot = 1000, method = "knn")


```{r}
# Regular bootstrap
indexes <- createDataPartition(tran_final$BPMbin2, p = .8, list = F)

train <- tran_final[indexes, ]
test <- tran_final[-indexes, ]

train_x <- train[,c(-2,-5,-6,-7,-8,-18,-19,-21,-22,-23,-24,-25,-36)]
train_y <- train[,36]

test_x <- test[,c(-2,-5,-6,-7,-8,-18,-19,-21,-22,-23,-24,-25,-36)]
test_y <- test[,36]

knnmodel1 <- knnreg(train_x, train_y)
#str(knnmodel1)
pred_y1 <- predict(knnmodel1, data.frame(test_x))
mse <- mean((test_y - pred_y1)^2)
mae <- caret::MAE(test_y,pred_y1)
rmse <- caret::RMSE(test_y,pred_y1)
cat("KNN Model 1: BPM Prediction","MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)
```
This is from determined based on ALL the data... we'll do additional tests based on future relevant data... for the sake of simplicity, we will use the same partition regarding BPM testing from the previous season

Test without binary...

```{r}
# This model uses Year (Pre-transfer year), transfer year, conference rating (old), conf_rating2 (post-transfer), and BPM binary before transfer, to predict BPM binary for the season after transfer  
train_x2 <- train[,c(3,4,10,27,35)]
train_y2 <- train[,36]

test_x2 <- test[,c(3,4,10,27,35)]
test_y2 <- test[,36]

knnmodel11 <- knnreg(train_x2, train_y2)
pred_y11 <- predict(knnmodel11, data.frame(test_x2))
mse11 <- mean((test_y2 - pred_y11)^2)
mae11 <- caret::MAE(test_y2,pred_y11)
rmse11 <- caret::RMSE(test_y2,pred_y11)
cat("KNN Model 2: BPM Prediction","MSE: ", mse11, "MAE: ", mae11, " RMSE: ", rmse11)
```

Are there other models to consider? What would be the best way to do model selection here? 

```{r}
# Creating data sets with only the required values
train_stats <- train[,c(9:17)]
test_stats_x <- test[,c(9:17)]
train_y_k <- train[,26]
test_y_k <- test[,26]
output <- data.frame(matrix(ncol=4,nrow=19))
colnames(output) <- c('MSE','MAE','RMSE','k')
# for loop to print knnreg results based on k=2:10
for (i in 2:20){
  knnmodel = knnreg(train_stats,train_y_k,k=i)
  #cat(knnmodel,i)
  pred_y <- predict(knnmodel, data.frame(test_stats_x))
  #cat(pred_y,i)
  output[i-1,4] <- i
  output[i-1,1] <- mean((test_y_k - pred_y)^2)
  output[i-1,2] <- caret::MAE(test_y_k,pred_y)
  output[i-1,3] <- caret::RMSE(test_y_k,pred_y)
  cat("KNN Model", i, ": BPM Prediction","MSE: ", output[i-1,1], "MAE: ", output[i-1,2], " RMSE: ", output[i-1,3], '\n')
}
ggplot(data = output) +
  geom_point(aes(x = k, y= RMSE), color = 'red', show.legend = FALSE) + 
  geom_point(aes(x = k, y = MSE), color = 'blue', show.legend = FALSE) + 
  geom_point(aes(x = k, y = MAE), color = 'green', show.legend = FALSE) +
  xlab("k") +
  ylab("Error")
mins <- apply(output[,1:3], 2, min)
mins
which(output[,1] == mins[1])
which(output[,2] == mins[2])
which(output[,3] == mins[3])
# optimal k here is 5 or 6... let's see what it is when we change the number of variables
```

```{r}
# Creating data sets with only the required values
train_stats_2 <- train[,c(9,14,16)]
test_stats_x2 <- test[,c(9,14,16)]
output2 <- data.frame(matrix(ncol=4,nrow=19))
colnames(output2) <- c('MSE','MAE','RMSE','k')
# for loop to print knnreg results based on k=2:10
for (i in 2:20){
  knnmodel = knnreg(train_stats_2,train_y_k,k=i)
  #cat(knnmodel,i)
  pred_y <- predict(knnmodel, data.frame(test_stats_x2))
  #cat(pred_y,i)
  output2[i-1,4] <- i
  output2[i-1,1] <- mean((test_y_k - pred_y)^2)
  output2[i-1,2] <- caret::MAE(test_y_k,pred_y)
  output2[i-1,3] <- caret::RMSE(test_y_k,pred_y)
  cat("KNN Model", i, ": BPM Prediction","MSE: ", output2[i-1,1], "MAE: ", output2[i-1,2], " RMSE: ", output2[i-1,3], '\n')
}
ggplot(data = output2) +
  geom_point(aes(x = k, y= RMSE), color = 'red', show.legend = FALSE) + 
  geom_point(aes(x = k, y = MSE), color = 'blue', show.legend = FALSE) + 
  geom_point(aes(x = k, y = MAE), color = 'green', show.legend = FALSE) +
  xlab("k") +
  ylab("Error")
mins2 <- apply(output2[,1:3], 2, min)
which(output2[,1] == mins2[1])
which(output2[,2] == mins2[2])
which(output2[,3] == mins2[3])
# optimal k here is 16... which seems high for a knn analysis
```


```{r}
library(corrplot)
library(MASS)
corData <- tran_final[,c(9:17,26,35:36)]
corrplot(cor(corData[,c(1:9)]))
# Based on the correlation heatmap, eFG and TS_per are highly correlated, so we should take out one of them (eFG in this instance). Also removing conf_rating2 because they should all be 1... not sure why I haven't removed it previously
# For regular regression, we should "lose" the binary values... and just use the continuous values of BPM
dat1 <- corData[,c(-4,-11,-12)]
corrplot(cor(dat1[,1:8]))
full.fit <- lm(BPM2 ~ ., data = dat1)
summary(full.fit)
step.AIC <- stepAIC(full.fit, direction = 'both', k =2)

alpha.1 <- 0.1
alpha.2 <- 0.15
step.wise <- olsrr::ols_step_both_p(full.fit, pent = alpha.1, prem = alpha.2)

step.wise$predictors  #No indus, exactly as we would expect
mod2 <- lm(BPM2 ~ BPM + blk_per + TS_per, data = dat1)
AIC(mod2)
mod3 <- lm(BPM2 ~ BPM + usg + TS_per + blk_per + stl_per, data = dat1)
AIC(mod3)
mod4 <- lm(BPM2 ~ BPM + TS_per + blk_per + stl_per, data = dat1)
AIC(mod4)

cat(" 3 Variables (BPM, blk_per, TS_per):",'\t\t', AIC(mod2), '\n', "5 variables (including usg and stl_per):",'\t', AIC(mod3), '\n', "4 variables (including stl_per):",'\t\t', AIC(mod4))
```


```{r}
Best.subset <- olsrr::ols_step_best_subset(full.fit)
which.max(Best.subset$rsquare)
which.max(Best.subset$adjr)
predictors.include <- strsplit(Best.subset$predictors[which.max(Best.subset$adjr)], "[ ]+", perl=T)[[1]]
which.min(Best.subset$aic)
aic.include <- strsplit(Best.subset$predictors[which.min(Best.subset$aic)], "[ ]+", perl=T)[[1]]
colnames(dat1)[!(colnames(dat1)%in%predictors.include)]  #Which is left out?
colnames(dat1)[!(colnames(dat1)%in%aic.include)]  #Which is left out?

alpha.3 <- 0.1
backward.1 <- olsrr::ols_step_backward_p(full.fit, penter = alpha.3)
backward.1$removed    #Model contains all covariates, except for indus.
```

```{r}
# Creating data sets with only the required values
train_stats_2 <- train[,c(9,14,16)]
test_stats_x2 <- test[,c(9,14,16)]
output2 <- data.frame(matrix(ncol=4,nrow=19))
colnames(output2) <- c('MSE','MAE','RMSE','k')
# for loop to print knnreg results based on k=2:10
for (i in 2:20){
  knnmodel = knnreg(train_stats_2,train_y_k,k=i)
  #cat(knnmodel,i)
  pred_y <- predict(knnmodel, data.frame(test_stats_x2))
  #cat(pred_y,i)
  output2[i-1,4] <- i
  output2[i-1,1] <- mean((test_y_k - pred_y)^2)
  output2[i-1,2] <- caret::MAE(test_y_k,pred_y)
  output2[i-1,3] <- caret::RMSE(test_y_k,pred_y)
  cat("KNN Model", i, ": BPM Prediction","MSE: ", output2[i-1,1], "MAE: ", output2[i-1,2], " RMSE: ", output2[i-1,3], '\n')
}
ggplot(data = output2) +
  geom_point(aes(x = k, y= RMSE), color = 'red', show.legend = FALSE) + 
  geom_point(aes(x = k, y = MSE), color = 'blue', show.legend = FALSE) + 
  geom_point(aes(x = k, y = MAE), color = 'green', show.legend = FALSE) +
  xlab("k") +
  ylab("Error")
mins2 <- apply(output2[,1:3], 2, min)
which(output2[,1] == mins2[1])
which(output2[,2] == mins2[2])
which(output2[,3] == mins2[3])
# optimal k here is 16... which seems high for a knn analysis
```

```{r}
# Logistic regression variable selection
l_data <- tran_final[,c(10:17,35,36)]
full.log.model <- glm(BPMbin2 ~ ., data = l_data, family = binomial)
step.log.model <- full.log.model %>% stepAIC(trace = FALSE)
coef(step.log.model)
```
Super interesting -- based on the above stepwise selection, the only 2 variables that are predictors to whether a player has a positive or negative BPM after transfer is based on the previous year and their blocks per game... curious why.
```{r}
# Creating data sets with only the required values
train_stats_3 <- train[,c(26,35)]
test_stats_x3 <- test[,c(16,35)]
output3 <- data.frame(matrix(ncol=4,nrow=19))
colnames(output3) <- c('MSE','MAE','RMSE','k')
# for loop to print knnreg results based on k=2:20
for (i in 2:20){
  knnmodel = knnreg(train_stats_3,train_y,k=i)
  #cat(knnmodel,i)
  pred_y <- predict(knnmodel, data.frame(test_stats_x3))
  #cat(pred_y,i)
  output3[i-1,4] <- i
  output3[i-1,1] <- mean((test_y - pred_y)^2)
  output3[i-1,2] <- caret::MAE(test_y,pred_y)
  output3[i-1,3] <- caret::RMSE(test_y,pred_y)
  cat("KNN Model", i, ": BPM Prediction","MSE: ", output3[i-1,1], "MAE: ", output3[i-1,2], " RMSE: ", output3[i-1,3], '\n')
}
ggplot(data = output3) +
  geom_point(aes(x = k, y= RMSE), color = 'red', show.legend = FALSE) + 
  geom_point(aes(x = k, y = MSE), color = 'blue', show.legend = FALSE) + 
  geom_point(aes(x = k, y = MAE), color = 'green', show.legend = FALSE) +
  xlab("k") +
  ylab("Error")
mins3 <- apply(output3[,1:3], 2, min)
which(output3[,1] == mins3[1])
which(output3[,2] == mins3[2])
which(output3[,3] == mins3[3])
# optimal k here is 3 or 14... which seems either high or low (although 3 is closer within the realm or normal) for a knn analysis
```

