---
title: "Transfer data"
author: "Nathan Matten"
date: "3/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(shipunov)
library(caret)
```

```{r}
acc_2017 <- read.csv(file = "ACC_2017_22.csv", header = T)
big12_2017 <- read.csv(file = "Big12_2017_22.csv", header = T)
big10_2017 <- read.csv(file = "Big10_2017_22.csv", header = T)
bigeast_2017 <- read.csv(file = "BigEast_2017_22.csv", header = T)
sec_2017 <- read.csv(file = "SEC_2017_22.csv", header = T)
pac12_2017 <- read.csv(file = "Pac12_2017_22.csv", header = T)
tran_2015 <- read.csv(file = "trank_data_15.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2015$player = paste(tran_2015$player_name, "(15)")
tran_2016 <- read.csv(file = "trank_data_16.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2016$player = paste(tran_2016$player_name, "(16)")
tran_2017 <- read.csv(file = "trank_data_17.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2017$player = paste(tran_2017$player_name, "(17)")
tran_2018 <- read.csv(file = "trank_data_18.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2018$player = paste(tran_2018$player_name, "(18)")
tran_2019 <- read.csv(file = "trank_data_19.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2019$player = paste(tran_2019$player_name, "(19)")
tran_2020 <- read.csv(file = "trank_data_20.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2020$player = paste(tran_2020$player_name, "(20)")
tran_2021 <- read.csv(file = "trank_data_21.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2021$player = paste(tran_2021$player_name, "(21)")
tran_2022 <- read.csv(file = "trank_data_22.csv", header = F, col.names = c("player_name","team","conf","GP","Min_per","ORtg","usg","eFG","TS_per","ORB_per","DRB_per","AST_per","TO_per","FTM","FTA","FT_per","twoPM","twoPA","twoP_per","TPM","TPA","TP_per","blk_per","stl_per","ftr","yr","ht","num","porpag",NA,NA,"year","pid","type","Rec Rank","ast/tov",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,"BPM",NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
tran_2022$player = paste(tran_2022$player_name, "(22)")
cbb_teams <- read.csv(file = "cbb_teams.csv", header = T)
# Creating mega reference table
tran_stats <- rbind(tran_2015,tran_2016,tran_2017,tran_2018,tran_2019,tran_2020,tran_2021,tran_2022) %>% 
  select(player,GP,Min_per,ORtg,usg,eFG,TS_per,ORB_per,DRB_per,AST_per,TO_per,FTM,FTA,FT_per,twoPM,twoPA,twoP_per,TPM,TPA,TP_per,blk_per,stl_per,ftr,yr,porpag,pid,BPM,year)
``` 

```{r}
# Making 1 giant transfer file
tran <- rbind(big10_2017,big12_2017,bigeast_2017,acc_2017,sec_2017,pac12_2017) %>% 
  select(-RK) %>% 
  arrange(PLAYER)
# Getting conferences for each transfer (from and to)
tran_team <- 
  tibble(
    player = tran$PLAYER,
    year = tran$YEAR.1,
    old_team = tran$OLD.TEAM,
    new_team = tran$TEAM,
    grade = tran$YEAR,
   ) %>% 
  arrange(old_team, year)

# Mapping new teams
tran_team <-  merge(tran_team, cbb_teams, by.x = "old_team", by.y = "Team", all.x = TRUE) 
tran_team <-  merge(tran_team, cbb_teams, by.x = "new_team", by.y = "Team", all.x = TRUE, no.dups = TRUE) 

# Setting up and arranging new table
tran_team <- tran_team %>%
  mutate(old_conference = Conference.x) %>% 
  mutate(new_conference = Conference.y) %>% 
  select(player,old_team,old_conference,new_team,new_conference,grade,year) %>%
  # Used this to search for conferences that didn't match in the mapping file
  #filter(is.na(old_conference)) %>% 
  arrange(old_conference,old_team,year)
# adding in conference ratings
tran_team$conf_rating <- 
ifelse (tran_team$old_conference %in% c('ACC','SEC','Pac-12','Big 10','Big 12','Big East'),1,
  ifelse (tran_team$old_conference %in% c("AAC","MWC","Atlantic 10","WCC","MVC"),2,
    ifelse (tran_team$old_conference %in% c("MAC","C-USA","SoCon","CAA","Sun Belt","Ivy League"),3,
      ifelse (tran_team$old_conference %in% c("Summit League","WAC","Horizon","Big West","Big Sky","MAAC","OVC"),4,
        ifelse (tran_team$old_conference %in% c("Patriot","ASUN","Big South","America East"),5,
          ifelse (tran_team$old_conference %in% c("Southland","NEC"),6,7)
        )
      )
    )
  )
)
# Creating the large table of transfers and their stats based off the two tables (tran and tran_team above)
transfer <- merge(tran_team, tran, by.y = "PLAYER", by.x = "player", all.x = TRUE) %>% 
  arrange(old_conference,year,old_team,new_conference,new_team)
# Middle table to get P-ID to then go back and remap
t2 <- merge(tran_team, tran_stats, by.x = "player", by.y = "player") %>% 
  select(player,old_team,old_conference,new_team,new_conference,grade,year.x,pid,conf_rating) %>% 
  arrange(year.x,pid)
# Mapping based on P-ID... then creating the mega-table of desired data
tran_yay <- merge(t2,tran_stats, by.x = "pid", by.y = "pid") %>% 
  mutate(player = player.x) %>%
  mutate(tran_year = year.x) %>% 
  select(-player.y,-player.x,-year.x) %>% 
  arrange(old_conference,old_team,new_conference,new_team,player,year)
# Figuring out how to only pull the transfer year stats and the previously played year's stats
#tran_final <- tran_yay %>% 
#  filter((year == tran_year))# & row((year == tran_year)-1))
```
Creating finall tables to use for analysis
```{r}
#Using this as testing ground to get rows needed for data
rows <- which(tran_yay$year[] == tran_yay$tran_year[])
nrows <- rows - 1
frows <- c(rows,nrows)
tran_final <- tran_yay %>% 
  filter(row_number() %in% frows) %>% 
  select(pid,player,year,tran_year,old_conference,old_team,new_conference,new_team,BPM,conf_rating,usg,eFG,Min_per,TS_per,TP_per,blk_per,stl_per) %>% 
  arrange(old_conference,old_team,new_conference,new_team,player,year)
# That seemed to work! Have 877 unique rows now... curious if there was a better way to do it? 
tran_final$year <- as.numeric(tran_final$year)
```

```{r}
# Now running KNN Bootstrap... let's see how it goes
tran_sub <- which(tran_final$year != tran_final$tran_year)
tran_final$BPMbin <- ifelse(tran_final$BPM >= 0, 1, 0)
boots <- tran_final %>% 
  select(year,BPMbin,conf_rating)
kboots <- BootKNN(boots,classes = tran_final$BPMbin,sub = tran_sub, nboot = 1000, method = "knn")
```

```{r}
# Regular bootstrap
indexes <- createDataPartition(tran_final$BPMbin, p = .8, list = F)

train <- tran_final[indexes, ]
test <- tran_final[-indexes, ]

train_x <- train[,c(-2,-5,-6,-7,-8,-18)]
train_y <- train[,18]

test_x <- test[,c(-2,-5,-6,-7,-8,-18)]
test_y <- test[,18]

knnmodel1 <- knnreg(train_x, train_y)
#str(knnmodel1)
pred_y1 <- predict(knnmodel1, data.frame(test_x))
mse <- mean((test_y - pred_y1)^2)
mae <- caret::MAE(test_y,pred_y1)
rmse <- caret::RMSE(test_y,pred_y1)
cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)
```
So my biggest question is -- this was not done correctly because the data is paired, but the partition is NOT paired. I'm not sure currently how to better pair it other than making duplicate columns for each player and the different years they played and doing this that way. Will follow up with you tomorrow during our discussion.



knn - season before, bpm, conference_rank (maybe position) - have to create the data for season before... might have a bit of an issue there, but gonna try on Sunday night.

turn BPM into binary, positive = 1, negative = 0...

